Challenges and Solutions of Developing and Implementing a Revolutionary Novel Desktop-as-a-Service

### 2024-04-21 paper.tex
Added related work (section 2) and a big part of architecture (section 3)

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Revolutionary Novel Desktop-as-a-Service}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun\orcidID{0009-0004-9955-3752} \and
	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering, Frankfurt University of Applied Sciences,\\
	Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
	\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	The paper will cover very novel and non-published knowledge we gained
	from developing and implementing a DaaS that is funded by the German
	Federal Ministry for Economic Affairs and Climate Action. The topic
	fits very well into the topics of the conference. TODO: Christian

@@

\section{Related Work}
\label{sec:relatedworkAbschnitt}
% TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
TODO: Christian (1)
TODO: Johannes (2)

\section{Architecture}
\label{sec:RecommendDaaSarchitecture}

(see Figure~\ref{figure_architecture}).

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Revolutionary Novel Desktop-as-a-Service}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun\orcidID{0009-0004-9955-3752} \and
	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
Frankfurt University of Applied Sciences,\\
Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
	The paper will cover very novel and non-published knowledge we gained
	from developing and implementing a DaaS that is funded by the German
	Federal Ministry for Economic Affairs and Climate Action. The topic
	fits very well into the topics of the conference. TODO: Christian

@@

\section{Related Work}
\label{sec:relatedworkAbschnitt}
% TODO:  Introduce: Infratstructure-as-a-Service  and others (IaaS, PaaS, SaaS, DaaS)
% TODO: Christian (1)
% TODO: Johannes (2)

DaaS has been a well-known service category since the emergence of cloud computing. Still, it got much less attention in research and literature
The challenges and obstacles when developing DaaS solutions, rather than infrastructure (IaaS) and platform services (PaaS), have seldom been discussed in the literature.

Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser. One focus of the paper is the redirection of the sound interface of the virtual desktop. The paper's authors conclude that Guacamole, in combination with the protocol RDP, is the best solution. The authors evaluated the lag time for remote audio playback in an Internet scenario. by measuring the time between the automated initiation of the live streaming using the VLC media player software and the arrival of the music data by analyzing network traffic via monitoring software Wireshark. When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms. When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms. The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application. Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms. The configuration of Guacamole and VNC caused an average lag of around 500\,ms.

Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations. The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, Microsoft RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA). In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming. One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios. In the Office Software Suite scenario, the transfer rate is measured and compared when performing several different tasks like opening and editing a document (typing), loading an image, and saving a document. The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions. The authors conclude that RDP requires fewer downstream and upstream network resources for all three scenarios. It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems. Depending on the different quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.

Several works in literature have analyzed the latency of Cloud gaming services that have, in some aspects, similar requirements to DaaS offerings. For both cloud service categories, users expect a latency value that is not considered annoying when interacting with the service and its applications. Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms. This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance.

\section{Architecture}
\label{sec:RecommendDaaSarchitecture}
% 
% TODO: Christian (1)
% TODO: Johannes (2)

Figure~\ref{figure_architecture} shows the architecture of our novel DaaS solution DESIGN that can run unmodified Linux and Windows applications in Linux containers (Linux and Windows applications compatible with Wine) and virtual machines (Windows applications incompatible with Wine).

The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container. Since the API of LXC is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution. A deployment of docker directly in the host-operating system or inside a virtual machine is both possible.

Exporting the graphical user interface of the Linux and Windows applications,
is possible using the a protocols VNC or RDP. Free server implementations exist for both operating system families. Since the focus of our DaaS solution is to export only the application's graphical user interface and not full desktops -- a feature not all implementations include -- is the number of available projects and solutions limited. 
For Linux containers, the free software projects x11vnc and Xrdp can be used in the DESIGN DaaS. For Windows VMs, the free software project TightVNC and the Windows-internal RDP server can be used.

One of the main goals of DESIGN is to allow all interactions to be carried out solely by using a browser, a broker or proxy software that mediates between the VNC or RDP implementation and the browser is required. The most advanced solution available is the free software project Apache Guacamole because it supports VNC and RDP both and includes further relevant features regarding sound and printing.

The core component of DESIGN is our self-developed Web UNI and Control Application, that .... TODO!!!
